A Hierarchical Bayesian Framework for Inferring the Stellar Obliquity Distribution

### 2023-06-06 src/tex/ms.tex
[showyourwork] overleaf sync

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{physics}
\usepackage[super]{nth}
\usepackage{fontawesome}
\usepackage{hyperref}

\newcommand{\dfm}[1]{\textcolor{blue}{\textbf{From DFM:} #1}}
\newcommand{\edits}[1]{\textcolor{black}{#1}}

@@

To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
We test the following \edits{five} $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation, \edits{and lastly, a Beta distribution such that $\cos{\psi} \sim \mathrm{Beta}(3, 6)$ with the support being extended linearly from $[0, 1]$ to $[-1, 1]$}.
For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$. We assume the stellar spin axis is uniformly distributed around the planetary orbital axis in the azimuthal direction and the orbital inclination is $90\degr$.
The sampled $i_\star$ and $\lambda$ here are \emph{true} values. 
To simulate the observation process, we add Gaussian noise to the \emph{true} $\lambda$ and $i_\star$, using uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical of the literature sample \cite{Albrecht22}.
Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with both $\lambda$ and $i_\star$ likelihoods or $\lambda$ likelihood only.

@@

\begin{figure*}
    \script{polar_plot.py}
    \begin{centering}
        \includegraphics{figures/polar.pdf}
        \caption{Inferred stellar obliquity distribution for the subsample of exoplanetary systems with both sky-projected stellar obliquity measurements and stellar inclination measurements. Inference made purely from measured projected stellar obliquity and inference made from both measured projected stellar obliquity and stellar inclination are shown in blue and orange curves, respectively. The line and shallow contours represent the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
        \label{fig:polar}
    \end{centering}
\end{figure*}

>>>>

% Define document class
\documentclass[twocolumn,times]{aastex631}
\usepackage{showyourwork}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{physics}
\usepackage[super]{nth}
\usepackage{fontawesome5}
\usepackage{hyperref}

\newcommand{\edits}[1]{\textcolor{purple}{#1}}

@@

To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
We test the following \edits{five} $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, \edits{a Beta distribution such that $\cos{\psi} \sim \mathrm{Beta}(3, 6)$ with the support being extended linearly from $[0, 1]$ to $[-1, 1]$}, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation.
For each $\cos{\psi}$ distribution, we randomly generate \edits{400} samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$. We assume the stellar spin axis is uniformly distributed around the planetary orbital axis in the azimuthal direction and the orbital inclination is $90\degr$.
The sampled $i_\star$ and $\lambda$ here are \emph{true} values. 
To simulate the observation process, we add Gaussian noise to the \emph{true} $\lambda$ and $i_\star$, using uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical of the literature sample \cite{Albrecht22}.
Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with both $\lambda$ and $i_\star$ likelihoods or $\lambda$ likelihood only.

@@

\begin{figure*}
    \script{polar_plot.py}
    \begin{centering}
        \includegraphics{figures/polar.pdf}
        \caption{Inferred stellar obliquity distribution for the full sample of exoplanetary systems with sky-projected stellar obliquity measurements (left panel) and the subsample of systems with both sky-projected stellar obliquity measurements and stellar inclination measurements (middle and right panels). The subsample inference made purely from measured projected stellar obliquity is shown in the middle panel, and the subsample inference made from both measured projected stellar obliquity and stellar inclination is shown in the right panel. The line and shallow contours represent the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions. The samples are extracted from \cite{Albrecht21} and \cite{Albrecht22}.}
        \label{fig:polar}
    \end{centering}
\end{figure*}
